---
title: "Putting your data on a map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Putting your data on a map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.2,
  dpi = 72,
  eval = rlang::is_installed("ggplot2")
)
```

```{r setup}
library(mongolmaps)
```

The usual workflow has three steps: get a table with one row per place,
join it to boundaries with `mn_join()`, and draw it with `mn_map()`.

## A small table

Place names can be written any common way, in English or Cyrillic:

```{r small}
cases <- data.frame(
  aimag = c("Khovsgol", "Hovd", "\u0423\u0432\u0441", "Ulan Bator", "Dornogobi"),
  cases = c(12, 30, 7, 140, 9)
)
cases_map <- mn_join(cases, by = aimag)
cases_map[c("name", "cases")]
```

Every aimag stays in the result, so places without data show up grey:

```{r small-map}
mn_map(cases_map, fill = cases)
```

## NSO tables

Tables from the National Statistics Office list several levels in one
column (the national total, regions, aimags, soums ...). Join them by the
code column, not the label column: labels such as "Ulaanbaatar" name both a
region and an aimag. `mn_join()` keeps the level you ask for and drops the
rest with a message.

```{r nso}
head(mn_example_population)

pop_2025 <- mn_example_population[mn_example_population$Year == 2025, ]
aimag_pop <- mn_join(pop_2025, by = "Region", level = "aimag")
mn_map(aimag_pop, fill = value / area_km2, title = "People per km2, 2025")
```

The same table has soum figures:

```{r soums}
soum_pop <- mn_join(pop_2025, by = "Region", level = "soum")
mn_map(soum_pop, fill = log10(value), title = "Soum population (log10), 2025")
```

## Several rows per place

Rows for several years give several copies of each polygon, ready for
facets:

```{r facets, fig.height = 6}
pop_years <- mn_join(mn_example_population, by = "Region", level = "aimag")
mn_map(pop_years, fill = value / 1000) + ggplot2::facet_wrap(~Year, ncol = 2)
```

## Repeated soum names

Many soums share a name. Give each row's aimag with `by_parent`:

```{r parent}
soums <- data.frame(
  aimag = c("Dornod", "Govi-Altai", "Khentii"),
  soum = c("Bayan-Uul", "Bayan-Uul", "Bayan-Adarga"),
  herders = c(820, 640, 910)
)
joined <- mn_join(soums, by = "soum", level = "soum", by_parent = "aimag")
joined[!is.na(joined$herders), c("name", "aimag_pcode", "herders")]
```

## Fetching NSO data with mongolstats

The mongolstats package downloads any NSO table. Its `Region` codes work
directly with `mn_join()`:

```{r mongolstats, eval = FALSE}
library(mongolstats)
tbl <- "DT_NSO_0300_002V4"
regions <- nso_dim_values(tbl, "Region")$code
years <- nso_dim_values(tbl, "Year", labels = "en")
latest <- years$code[1]
pop <- nso_data(tbl, selections = list(Region = regions, Year = latest), labels = "en")
mn_map(mn_join(pop, by = "Region", level = "aimag"), fill = value)
```

## Checking the matches

`mn_match()` shows what each value matches, and warns about anything
ambiguous or unmatched:

```{r check}
mn_match(c("Khovd", "Hovd", "Kobdo", "Jargalant"), to = "name_en")
```
