---
title: "Getting started with mongolmaps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with mongolmaps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.2,
  dpi = 72,
  eval = rlang::is_installed("ggplot2")
)
```

mongolmaps gives you ready-made maps of Mongolia. Everything in this
vignette works offline: simplified boundaries ship with the package.

```{r setup}
library(mongolmaps)
```

## Administrative levels

Mongolia has 21 aimags (provinces) and the capital, Ulaanbaatar. Aimags are
divided into soums, and soums into bags. Ulaanbaatar is divided into 9
districts (duureg) and those into khoroos.

| Level in mongolmaps | Rural units | Ulaanbaatar units |
|---|---|---|
| `"country"` | Mongolia | |
| `"region"` | 4 economic regions | Ulaanbaatar is the 5th |
| `"aimag"` | 21 aimags | Ulaanbaatar |
| `"soum"` | 330 soums | 9 districts |
| `"bag"` | ~1,650 bags | 204 khoroos |

Each level has a function:

```{r levels}
mn_aimags()
```

All of them return the same columns:

* `pcode`: a code that identifies the unit at any level;
* `name`, `name_en`, `name_mn`, `name_mns`: the name in the chosen
  language, in English (as NSO writes it), in Cyrillic, and in Latin with
  diacritics;
* `level`, `type`, `number`: what kind of unit it is;
* `iso_code`, `nso_code`: ISO 3166-2 and NSO statistical codes;
* `parent_pcode`, `region_pcode`, `aimag_pcode`, `soum_pcode`: the units it
  sits in;
* `area_km2` and the `geometry`.

## Filtering

Pass the places you want. Names can be spelled any common way.

```{r filter}
mn_soums(aimag = "Khovd")
mn_aimags(region = "Western")
```

## Drawing maps

`mn_map()` draws any of these with ggplot2.

```{r map-soums}
mn_map(mn_soums(aimag = c("Khovd", "Uvs", "Bayan-Ulgii")), fill = aimag_pcode, label = TRUE)
```

It returns a normal ggplot, so you can keep adding to it:

```{r map-add}
library(ggplot2)
mn_map(mn_aimags()) +
  geom_sf(data = mn_settlements(type = c("capital", "aimag_centre")), colour = "firebrick") +
  labs(title = "Aimag centres")
```

With `context = TRUE` you get neighbouring countries, rivers and lakes:

```{r map-context}
mn_map(mn_regions(), fill = name, context = TRUE)
```

## Projections

Boundaries come in longitude and latitude (EPSG:4326). `mn_map()` projects
them for you; to project them yourself use `crs`:

```{r crs}
mn_aimags(crs = "albers")
```

`mn_crs()` returns the projections: Albers equal-area for national maps,
UTM zone 48N for Ulaanbaatar.

## Languages

```{r mongolian}
mn_map(mn_aimags(lang = "mn"), label = TRUE, lang = "mn")
```

Set `options(mongolmaps.lang = "mn")` to use Mongolian everywhere.

## Higher resolution

The bundled boundaries are simplified. For detailed work, use
`resolution = "high"`: the full-resolution boundaries (about 10 MB) are
downloaded once and cached.

```{r high, eval = FALSE}
mn_soums(aimag = "Khovd", resolution = "high")
```

## Where the data come from

```{r sources}
mn_sources()[c("id", "provider", "license")]
```

Please credit the providers; `mn_citation()` gives the text.
