---
title: "Bags: using boundaries from NSO"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(mongolmaps)
```

Bags (baga) are the smallest rural units: each soum is divided into a few
bags, about 1,650 in all. mongolmaps knows every bag's code and name:

```{r codes}
mn_codes("bag", within = "Khovd")
```

but **not their boundaries**. No open dataset publishes them: GADM,
geoBoundaries, the humanitarian COD and OpenStreetMap all stop at soums. The
National Statistics Office of Mongolia (NSO) shares bag boundaries on
request.

## Getting the file

Write to NSO at <international@nso.mn> and ask for:

* bag boundaries as a shapefile or GeoPackage;
* each bag's NSO code (7 digits, such as `1830151`) or its soum code and
  bag number;
* the date the boundaries are valid for;
* the terms of use (can you share maps? can the file be redistributed?).

## Using the file

Point mongolmaps at the file:

```{r use, eval = FALSE}
bags <- mn_bags(path = "bags_from_nso.gpkg")
mn_map(mn_bags(aimag = "Khovd", path = "bags_from_nso.gpkg"), label = TRUE)
```

or register it once, for example in your `.Rprofile`, and forget about it:

```{r register, eval = FALSE}
options(mongolmaps.bags_path = "C:/data/bags_from_nso.gpkg")

mn_bags(aimag = "Khovd")
mn_admin("bag") # khoroos and bags together
mn_join(my_bag_table, by = "code", level = "bag")
```

`mn_read_bags()` checks the file when it is first read:

* it finds the column of bag codes (or use `code_col`);
* every code must be a known NSO bag code;
* every bag must lie inside its soum (turn off with
  `check_nesting = FALSE`).

Bags then get the same columns as every other map in the package, so
joining NSO bag tables works like joining aimag tables.
