---
title: "Terrain, land cover, population and protected areas"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  eval = identical(Sys.getenv("NOT_CRAN"), "true") && rlang::is_installed("terra")
)
```

```{r setup}
library(mongolmaps)
library(ggplot2)
```

These layers are rasters (grids of cells), returned as `terra` SpatRasters.
National 1 km grids are downloaded once and cached; finer grids are read
directly from the cloud for one area at a time.

| Function | Resolutions | Source |
|---|---|---|
| `mn_elevation()`, `mn_hillshade()` | 1 km (national), 90 m (by area) | Copernicus DEM GLO-90 |
| `mn_landcover()` | 1 km (national), 10 m (by area) | ESA WorldCover 2021 |
| `mn_population()` | 1 km, 100 m; years 2015 to 2030 | WorldPop |

`mn_map()` draws rasters too, with the right colours and credits.

## Terrain

```{r elevation}
mn_map(mn_elevation(), title = "Elevation (m)") +
  geom_sf(data = mn_aimags(), fill = NA, colour = "white", linewidth = 0.2)
```

Shaded relief makes a good background under other layers:

```{r hillshade}
mn_map(mn_hillshade(), caption = FALSE) +
  geom_sf(data = mn_aimags(), fill = NA, colour = "white", linewidth = 0.3) +
  geom_sf(data = mn_settlements(type = c("capital", "aimag_centre")), colour = "firebrick", size = 1)
```

For one area, 90 m detail:

```{r bogd}
mn_map(mn_hillshade("90m", within = "Ulaanbaatar"), caption = FALSE) +
  geom_sf(data = mn_ub_districts(), fill = NA, colour = "firebrick")
```

## Land cover

```{r landcover}
mn_map(mn_landcover(), title = "Land cover, 2021")
```

## Population

`mn_zonal()` adds up the grid for any set of polygons, for example soums:

```{r population}
pop <- mn_population(2025)
soums <- mn_zonal(pop, mn_soums(), name = "population")
mn_map(soums, fill = population / area_km2, trans = "log10",
       title = "People per km2 by soum, 2025 (WorldPop)")
```

## Protected areas

The World Database on Protected Areas may not be redistributed, so it is
downloaded from UNEP-WCMC on first use:

```{r protected}
pa <- mn_protected_areas()
mn_map(mn_aimags(), caption = FALSE) +
  geom_sf(data = pa, aes(fill = desig_eng), alpha = 0.6, colour = NA) +
  labs(fill = NULL, caption = mn_citation(c("admin", "wdpa")))
```
