---
title: "Thematic layers: roads, water, places"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  eval = identical(Sys.getenv("NOT_CRAN"), "true")
)
```

```{r setup}
library(mongolmaps)
library(ggplot2)
```

Besides boundaries, mongolmaps provides the layers that make a map
readable. Small ones ship with the package; larger ones are downloaded once
and cached (see `mn_cache_dir()`).

| Function | What | Source | Delivery |
|---|---|---|---|
| `mn_settlements()` | capital, aimag and soum centres | Wikidata | bundled |
| `mn_neighbours()` | Russia, China, Kazakhstan around Mongolia | Natural Earth | bundled |
| `mn_rivers()`, `mn_lakes()` | major rivers and lakes | Natural Earth | bundled |
| `mn_rivers("all")`, `mn_lakes("all")` | every river, stream and water body | OpenStreetMap | download, ~20 MB |
| `mn_roads()` | roads and streets | OpenStreetMap | download, ~30 MB |
| `mn_railways()`, `mn_airports()` | railways, stations, airports | OpenStreetMap | download, <1 MB |
| `mn_places()` | cities, towns, villages | OpenStreetMap | download, <1 MB |

## The country

```{r country}
mn_map(mn_aimags(), context = TRUE) +
  geom_sf(data = mn_roads(), colour = "firebrick", linewidth = 0.25) +
  geom_sf(data = mn_railways(), colour = "black", linewidth = 0.6) +
  geom_sf(data = mn_airports(), shape = 17, colour = "navy") +
  labs(title = "Main roads, railways and airports")
```

## One aimag

`within` keeps the features inside a place and cuts lines at its border:

```{r khuvsgul, fig.height = 6}
aimag <- "Khuvsgul"
mn_map(mn_soums(aimag = aimag)) +
  geom_sf(data = mn_lakes("all", within = aimag), fill = "#9ecae1", colour = NA) +
  geom_sf(data = mn_rivers("all", within = aimag), colour = "#4292c6", linewidth = 0.2) +
  geom_sf(data = mn_roads(within = aimag), colour = "grey30", linewidth = 0.3) +
  geom_sf(data = mn_settlements(within = aimag), size = 1) +
  labs(title = "Khuvsgul: soums, water, roads and soum centres")
```

## Ulaanbaatar streets

```{r ub, fig.height = 6}
central <- c("BGD", "BZD", "ChD", "KhUD", "SKhD", "SBD")
streets <- mn_roads(class = c("main", "minor"), within = central)
mn_map(mn_khoroos(district = central), fill = parent_pcode) +
  geom_sf(data = streets, aes(linewidth = class), colour = "grey20") +
  scale_linewidth_manual(values = c(main = 0.5, minor = 0.1), guide = "none")
```

## Credits

OpenStreetMap layers are under the Open Database Licence. Credit them when
you publish:

```{r credits}
mn_citation(c("admin", "osm_roads", "osm_water"))
```
